---
title: "Technical Note 02"
author: "Roy Ratcliffe"
abstract: |
  What version of Bash does Linux normally equip? The short answer is version 5, typically.
output:
  pdf_document: pinp::pinp
  html_document:
    df_print: paged
one_column: true
---

```{r setup, include=FALSE}
library(docknitr)

knitr::opts_chunk$set(
  comment = NULL
)
```

# Introduction
This technical exploration _only_ considers two Linux distributions: Debian and Alpine. Debian subsumes Ubuntu; the latter adds a user-friendly graphical interface to the base Debian distribution for end-users, where the former primarily targets the server-side deployment.

# Debian equips Bash version 5
Version 5 for Debian version 11, short codename `bullseye`. No pre-installation required here because Debian sets up Bash as the default shell.
```{r debian-11, cache=TRUE, engine="docker", image="debian:11"}
echo $BASH_VERSION
```

# Alpine also version 5
Alpine's default shell is BusyBox, a multi-call binary. Add Bash first before asking its version.
```{r alpine-latest, cache=TRUE, engine="docker", image="alpine:latest"}
apk add bash
bash --version
```

# Further work
The investigation initially ignores other less-ubiquitous distributions of Linux. CentOS and other flavours leave scope for furthur work.
